---
title: "ADHD-SMART-operating-characteristics"
output: 
  rmarkdown::html_document:
    toc: TRUE
    fig-width: 100%
vignette: >
  %\VignetteIndexEntry{ADHD-SMART-operating-characteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(CATIE2023)
library(dplyr)
library(ggplot2)
library(tidyr)
library(geepack)
library(emmeans)

```

# Operating Characteristics Summary

For EOS outcome

-   First-stage:
    -   BMOD is marginally better

    -   IF Prior Med, MED is better (see 1st stage interaction plot)
-   Second-stage (among non-responders only):
    -   AUG is marginally better

    -   IF a1 = MED, INT is better; IF a1 = BMOD, AUG is better

    -   deeply tailored: IF Non-adherent, AUG always better, likewise, IF adherent, INT is better (see 2nd stage interaction plot)
-   Associations:
    -   Baseline: Odd, Severity neg assoc with Y2; Race pos associated with Y2

    -   Intermediate:

        -   Responder pos assoc with Y2

        -   Response is pos associated with MED, and neg associated with BMOD IF Prior Med

        -   Adherence is pos associated with MED IF Prior Med

        -   Non-response time (ET) is uniform random
-   Y0 neg associated with Odd and Severity
-   Y1 slight increase from baseline; pos accociated with MED only

## Linear DGP Model

$$
Y_{0,i}(a_1, a_2) = \eta_0 - \eta_1odd_{c} - \eta_2 severity_c + \epsilon_i
$$

$$
Y_{1,i}(a_1, a_2) = \eta_0 - \beta_1a_1 + \eta_2U_i + \eta_3Y_{0c}\epsilon_i
$$

$$
Y_{2,i}(a_1, a_2) = \eta_0 - \eta_1odd_c - \eta_2severity_c + \eta_3priormed_c + \eta_4race_c + \\
(\beta_1 - \beta_2priormed_c)a_1 + \\
\eta_5\tilde{R} + \eta_6 \tilde{Adh} + \eta_7\tilde{Y_1} \\
+(1-R)(\beta_3 + \beta_4a_1 + \beta_5Adh_c)a_2 + 0.4U + \epsilon_i
$$

# Import Data
```{r}

adhd <- CATIE2023::adhd
trueDTRmeans <- attr(adhd, "dtrmeans")

# Fill NA with zero
adhd[adhd$R == 1, "NRtime"] <- 0
adhd[adhd$R == 1, "A2"] <- 0
```

```{r}
# Manipulate data.frame

# Grand mean center covariates
adhd.c <- adhd %>% mutate(across(c(2:5, adherence, NRtime),  ~ .x - mean(.x, na.rm = TRUE), .names = "{.col}.c"), )

# Center variables only among non-responders
adhd.nr.c <- adhd %>% filter(R == 0) %>%
  mutate(across(.cols = c(2:5, NRtime, adherence), ~ .x - mean(.x), .names = "{.col}.c"))

# Long format for plotting longitudinal outcome
adhd.long <- pivot_longer(adhd, cols = c(Y0, Y1, Y2), names_to = "stage", 
                          names_prefix = "[Y]",
                          names_transform = as.integer,
                          values_to = "Y")

```

## Some Plots

```{r}
adhd.long %>% ggplot(aes(x = factor(stage), y = Y, fill =  factor(stage))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "Longitudinal Outcome", x = "Stage", y = "School Perfomance") +
  theme_bw()

adhd.long %>% ggplot(aes(x = factor(stage), y = Y, group = ID)) +
  geom_line(alpha = 0.2, color = "blue") +
  labs(title = "Individual Longitudinal Outcome", x = "Stage", y = "School Perfomance") +
  theme_bw()

adhd.long %>% filter(stage == 2) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  interaction(A1, A2, sep = " "))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "EOS Outcome by Cell", x = "Stage", y = "School Perfomance") +
  scale_fill_discrete("A1, A2") +
  theme_bw()

```

## Correlation between Outcomes
```{r}

adhd %>% select(Y0, Y1, Y2) %>% cor()

```



# Module 4a / DEMO 1

Main Effects

### Response Rate

```{r}
xtabs(~ R, adhd) %>% prop.table()
xtabs( ~ A1 + R, adhd) %>% prop.table()
xtabs( ~ R + A2, adhd) %>% prop.table()

# test for Responder difference
chisq.test(xtabs( ~ A1 + R, adhd), correct = FALSE)

modelR <- adhd.c %>% glm(R ~ A1 + odd.c + severity.c + priormed.c + race.c, 
                         family = binomial, data = .)
summary(modelR)

```

### First-stage Main Effects

```{r fig.cap="Longitudinal Outcome by first-stage treatment"}
# plot
adhd.long %>% filter(stage > 0) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  factor(A1))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "Longitudinal Outcome by First-Stage Tx", x = "Stage", y = "School Perfomance") +
  scale_fill_discrete("A1", labels = c("-1 MED", "1 BMOD"))
```

```{r}
# Fit model


model1 <- adhd.c %>%geepack::geeglm(Y2 ~ A1 + odd.c + severity.c + priormed.c + race.c, 
                            id = ID, data = .)
summary(model1)

estimate(model1,
         rbind("Mean Y under BMOD"   = c(1,  1, 0, 0, 0, 0),
               "Mean Y under MED"    = c(1, -1, 0, 0, 0, 0),
               "Between groups diff" = c(0,  2, 0, 0, 0, 0))) %>% print()

```

### Second-stage Among Non-responders

```{r}
# Plots of second stage
adhd.long %>% filter(stage == 2) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  factor(A2))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "EOS Outcome by Second-Stage Tx", x = "Stage", y = "School Perfomance") +
  scale_fill_discrete("A2", labels = c("-1 AUG", "Responder", "1 INT"))

```

```{r}

# Covariates centered among non-responders
model2 <- adhd.nr.c %>%geepack::geeglm(Y2 ~ A2 + odd.c + severity.c + priormed.c + 
                                 race.c + NRtime.c + adherence.c, id = ID, data = .)
summary(model2)

estimate(model2,
         rbind("Mean Y w/INTENSIFY tactic"   = c(1,  1, rep(0, 6)),
               "Mean Y w/AUGMENT tactic"     = c(1, -1, rep(0, 6)),
               "Between groups difference"   = c(0,  2, rep(0, 6)))) %>% print()
```

# Module 4b / DEMO 2

Compare all 4 AI using replication and weighting

### Replicate and Weight

```{r}
# Replicate Responders, add A2 tx, bind non-responders
# Use centered df
adhd.rep <- adhd.c %>% filter(R == 1) %>% 
  slice(rep(1:n(), each = 2)) %>% # Replicates
  mutate(A2 = rep(c(-1, 1), length.out = n()))

adhd.rep <- bind_rows(adhd.c %>% filter(R == 0))

# Add weights
adhd.rep <- adhd.rep %>% mutate(w = if_else(R == 1, 2, 4))


```

### Fit Marginal Model

```{r}
modelMM <-geepack::geeglm(Y2 ~ A1*A2, weights = w, id = ID, data = adhd.rep)
summary(modelMM)

# Careful we specify contrasts correctly (coeff in right order)
estimate(modelMM,
         rbind(
           # These statements get the mean under each AI
           "Mean Y: AI#1 (MED, AUGMENT)"    = c(1, -1, -1,  1),
           "Mean Y: AI#2 (BMOD, AUGMENT)"   = c(1,  1, -1, -1),
           "Mean Y: AI#3 (MED, INTENSIFY)"  = c(1, -1,  1, -1),
           "Mean Y: AI#4 (BMOD, INTENSIFY)" = c(1,  1,  1,  1),
           # These statements are to get all pairwise comparisons
           "Diff: AI#1 - AI#2"              = c(0, -2,  0,  2),
           "Diff: AI#1 - AI#3"              = c(0,  0, -2,  2),
           "Diff: AI#1 - AI#4"              = c(0, -2, -2,  0),
           "Diff: AI#2 - AI#3"              = c(0,  2, -2,  0),
           "Diff: AI#2 - AI#4"              = c(0,  0, -2, -2),
           "Diff: AI#3 - AI#4"              = c(0, -2,  0, -2)))

```


### Marginal Model, adjust for baseline covariates

```{r}
modelMM.adj <- adhd.rep %>%geepack::geeglm(Y2 ~ A1*A2 + odd.c + severity.c + priormed.c + race.c, 
                                   weights = w, id = ID, data = .)

summary(modelMM.adj)

# Careful we specify contrasts correctly
estimate(modelMM.adj,
         rbind("Mean Y: AI#1 (MED, AUGMENT)"    = c(1, -1, -1, rep(0,4),  1),
               "Mean Y: AI#2 (BMOD, AUGMENT)"   = c(1,  1, -1, rep(0,4), -1),
               "Mean Y: AI#3 (MED, INTENSIFY)"  = c(1, -1,  1, rep(0,4), -1),
               "Mean Y: AI#4 (BMOD, INTENSIFY)" = c(1,  1,  1, rep(0,4),  1),
               "Diff: AI#1 - AI#2"              = c(0, -2,  0, rep(0,4),  2),
               "Diff: AI#1 - AI#3"              = c(0,  0, -2, rep(0,4),  2),
               "Diff: AI#1 - AI#4"              = c(0, -2, -2, rep(0,4),  0),
               "Diff: AI#2 - AI#3"              = c(0,  2, -2, rep(0,4),  0),
               "Diff: AI#2 - AI#4"              = c(0,  0, -2, rep(0,4), -2),
               "Diff: AI#3 - AI#4"              = c(0, -2,  0, rep(0,4), -2))) %>% print()

```

# Module 5 / DEMO 3

### Moderators of Second-stage TX effects

First-stage effects will be biased Why do we subset to only non-responders? We loose power adjusting for baseline and first-stage?

```{r}
# should we use the df with the coeff centered by non-responders?
# do we interact with the centered adherence.c?
model2s.c <- adhd.nr.c %>%geepack::geeglm(Y2 ~ odd.c + severity.c + priormed.c + race.c +
                                  NRtime.c + adherence.c + 
                                    A1*A2 + A2:adherence.c,
                                id = ID, data = .)
summary(model2s.c) #gives robust SE
```

### Interaction Plot 2nd Stage

```{r}


model2s <- adhd.nr.c %>%geepack::geeglm(Y2 ~ odd.c + severity.c + priormed.c + race.c +
                                  NRtime.c + adherence + 
                                  A1*A2 + A2:adherence,
                                id = ID, data = .)

grid <- emmeans::ref_grid(model2s, cov.keep = c("A1","A2", "adherence"))
grid

em2 <- emmeans::emmeans(grid, ~ A1 + A2 | adherence, weight = "equal") # This does not give a weighted grid!!!! use ref_grid or weights = "proportional"
summary(em2)
plot(em2)

ep2 <- emmeans::emmip(em2, A1 + A2 ~ adherence, style = "factor")
ep2$data %>% mutate(across(1:3, as.factor)) %>%
  ggplot(aes(xvar, yvar, color = A1, linetype = A2, group = tvar)) +
  geom_line() +
  geom_point() +
  scale_color_manual("A1", values = c("-1" = "red", "1" = "blue"), 
                     labels = c("-1" = "-1 MED", "1" = "1 BMOD")) +
  scale_linetype_manual("A2", values = c("-1" = 1,"1" = 3), 
                        labels = c("-1" = "-1 AUG", "1" = "1 INT")) +
  xlab("Adherence") +
  ylab("School Performance") +
  theme_bw()

```

### Q-learning

control for optimal second-stage tactic

```{r}
# Using full dataset, centered first, then filtered to non-responders
adhd.c.nr <- adhd.c %>% filter(R == 0)

model2s <-  adhd.c.nr %>%
 geepack::geeglm(Y2 ~ odd.c + severity.c + priormed.c + race.c + 
           NRtime + adherence + A1*A2 + A2:adherence, 
         id = ID, data = .)

adhd.c.nr.alt <- adhd.c.nr %>% mutate(A2 = -1 * A2)
adhd.c.nr.opt <- adhd.c.nr

adhd.c.nr.alt$Y2 <- predict(model2s, newdata = adhd.c.nr.alt) # predict response on counterfactual a2

for (i in 1:nrow(adhd.c.nr.alt)) {
  if (adhd.c.nr.alt$Y2[i] > adhd.c.nr.opt$Y2[i]) {
    adhd.c.nr.opt[i, ] <- adhd.c.nr.alt[i, ] # assign optimal A2
  }
}

# Combine Optimal non-responders with responders
adhd.c.opt <- adhd.c %>% filter(R == 1) %>%
  bind_rows(adhd.c.nr.opt)

# Fit first-stage model controlling for second-stage
model1s <- adhd.c.opt %>%geepack::geeglm(Y2 ~ odd.c + severity.c + priormed.c + race.c +
                                   A1 + A1:priormed, id = ID, data = .)
summary(model1s)


```

### Interaction plot 1st Stage

```{r}
grid <- emmeans::ref_grid(model1s, cov.keep = c("A1","priormed"))
grid

em1 <- emmeans::emmeans(model1s, ~ A1 | priormed, weights = "equal")
summary(em1)

ep1 <- emmeans::emmip(em1, A1 ~ priormed, style = "factor")

ep1 + scale_color_manual("A1", values = c("-1" = "red", "1" = "blue"), 
                     labels = c("-1" = "-1 MED", "1" = "1 BMOD")) +
  labs(x = "Levels of Prior Med", y = "School Performance") +
  theme_bw()
```

TO DO: calculate mean of more deeply tailored AI


### qlaci

Need to implement qlaci package. Are we using?? I think there are some errors in the Q-learning demo...

### Adjust for Time-varying Confounders (conditional model)

Two-stage regression algorithm. SE errors need to be adjusted

```{r}

adhd.c$Rerand <- 1 - adhd.c$R

# Response model
fitR <- adhd.c %>% glm(R ~ odd.c + severity.c + priormed.c + race.c + A1 + adherence.c,
                       family = binomial, data = .)
#summary(fitR)

# Add residuals to df
adhd.c$R.resid <- residuals(fitR, "response")

# Adherence model
fitAdh <- adhd.c %>% glm(adherence ~  odd.c + severity.c + priormed.c + race.c + A1 + R,
                       family = binomial, data = .)
#summary(fitAdh)

# Add residuals to df
adhd.c$adherence.resid <- residuals(fitAdh, "response")

# Fit main model
fitCM <- adhd.c %>% lm(Y2 ~ A1*R.resid + Rerand:(A2/A1), 
                       data = .)

summary(fitCM)


```

### Moderator Analysis

```{r}

# Fit main model
fitCM.mod <- adhd.c %>% lm(Y2 ~ A1*(R.resid + odd.c + severity.c + priormed.c + race.c) + adherence.resid + Rerand:(A2/(A1 + adherence.c)), data = .)

summary(fitCM.mod)

```

