---
title: "AUTSIM-SMART-operating-characteristics"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{AUTSIM-SMART-operating-characteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(CATIE2023)
library(tidyverse)
```

## Linear DGP Model

$$
Y_{0,i}(a_1, a_2) = 40 + 0.4\bar{O}_{11} + \epsilon_i
$$

$$
Y_{1,i}(a_1, a_2) = 50 + 1a_1 + 4U_i + \epsilon_i
$$

$$
Y_{2,i}(a_1, a_2) = 60 + 0.3\bar{O}_{11} + (1 + 0.3\bar{O}_{11})a_1 + \\
18\tilde{R} + 5\tilde{O}_{21} \\
+I(a_1 =1)(1-R)(-4 + 6\tilde{O}_{21})a_2 + 6U_i + \epsilon_i
$$


# Import Data

```{r}
set.seed(2023-01-21)

# Generate autism
autismObj <- simAUTISMsmart(N=200) #original dataset N=200
autism.raw <- autismObj$data
trueDTRmeans <- autismObj$DTRmeans


# Fill NA with zero
autism <- autism.raw
autism[is.na(autism)] <- 0
```

```{r}
# Manipulate data.frame

autism <- autism %>% mutate(Rerand = (1 - R) * (A1 == 1))

# Grand mean center covariates
autism.c <- autism %>% mutate(across(c(O11, O12, O21, O22),  ~ .x - mean(.x), .names = "{.col}.c"))

# Center variables only among non-responders to A1==1
autism.nr.c <- autism %>% filter(R == 0, A1 == 1) %>%
  mutate(across(.cols = c(O11, O12, O21, O22), ~ .x - mean(.x), .names = "{.col}.c"))

# Long format for plotting longitudinal outcome
autism.long <- pivot_longer(autism, cols = c(Y0, Y1, Y2), names_to = "stage", 
                          names_prefix = "[Y]",
                          names_transform = as.integer,
                          values_to = "Y") %>%
  mutate(DTR = interaction(A1,A2, sep = " "))

```

## Some Plots

```{r}
autism.long %>% ggplot(aes(x = factor(stage), y = Y, fill =  factor(stage))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "Longitudinal Outcome", x = "Stage", y = "Score")

autism.long %>% filter(stage == 2) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  DTR)) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "EOS Outcome by Cell", x = "Stage", y = "Score") +
  scale_fill_discrete("A1, A2")

```

# Module 4a / DEMO 1

Main Effects

### Response Rate

```{r}
xtabs(~ R, autism) %>% prop.table()
xtabs( ~ A1 + R, autism) %>% prop.table(margin = 1)
xtabs( ~ R + A2, autism) %>% prop.table(margin = 1)

# test for Responder difference
chisq.test(xtabs( ~ A1 + R, autism), correct = FALSE)

modelR <- autism.c %>% glm(R ~ A1 + O11.c + O12.c, 
                         family = binomial, data = .)
summary(modelR)

```

### First-stage Main Effects

```{r fig.cap="Longitudinal Outcome by first-stage treatment"}
# plot
autism.long %>% filter(stage > 0) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  factor(A1))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "Longitudinal Outcome by First-Stage Tx", x = "Stage", y = "Score") +
  scale_fill_discrete("A1", labels = c("-1 SGD", "1 No SGD"))
```

```{r}
# Fit model
library(geepack)

model1 <- autism.c %>% geeglm(Y2 ~ A1 + O11.c + O12.c, 
                            id = ID, data = .)
summary(model1)

estimate(model1,
         rbind("Mean Y under NO SGD"   = c(1,  1, 0, 0),
               "Mean Y under SGD"    = c(1, -1, 0, 0),
               "Between groups diff" = c(0,  2, 0, 0)))

```

### Second-stage Among Non-responders

```{r}
# Plots of second stage
autism.long %>% filter(stage == 2, Rerand == 1) %>% 
  ggplot(aes(x = factor(stage), y = Y, fill =  factor(A2))) +
  geom_violin(draw_quantiles = 0.5) +
  labs(title = "EOS Outcome by Second-Stage Tx", x = "Stage", y = "Score") +
  scale_fill_discrete("A2", labels = c("-1 Add SGD", "1 INT"))

```

```{r}

# Covariates centered among non-responders
model2 <- autism.nr.c %>% geeglm(Y2 ~ A2 + O11.c + O12.c + O21.c + O22.c, id = ID, data = .)
summary(model2)

estimate(model2,
         rbind("Mean Y w/INTENSIFY tactic"   = c(1,  1, rep(0, 4)),
               "Mean Y w/AUGMENT tactic"     = c(1, -1, rep(0, 4)),
               "Between groups difference"   = c(0,  2, rep(0, 4))))
```

# Module 4b / DEMO 2

Compare all 4 AI using replication and weighting

### Replicate and Weight

```{r}
# Replicate Responders, add A2 tx, bind non-responders
# Use centered df
autism.rep <- autism.c %>% filter(A1 == 1, R == 1) %>%
  slice(rep(1:n(), each = 2)) %>% # Replicates
  mutate(A2 = rep(c(-1, 1), length.out = n()))

autism.mm <- autism.c %>% filter(A1 == -1 | R == 0) %>%
  bind_rows(autism.rep)

autism.mm <- autism.mm %>% mutate(w = if_else(Rerand == 1, 4, 2))


```

### Fit Marginal Model

```{r}
fitmm <- geepack::geeglm(Y2 ~ A1 + A1:A2, data = autism.mm, id = ID, weights = w)
summary(fitmm)

estimate(fitmm,
         rbind(
           # These statements get the mean under each AI
           # Remember the coefficients are positional: the first is the intercept, the second is the variable in the blank, the third is a1a2.
           "Mean Y: AI#1 (JASP+EMT,INTENSFY)" = c(1, 1, 1),
           "Mean Y: AI#2 (JASP+EMT, Add SGD)" = c(1, 1, -1),
           "Mean Y: AI#3 (JASP+EMT+SGD, ...)" = c(1, -1, 0),
           "Diff: (JASP+EMT,Add SGD) - (JASP+EMT+SGD,...)" = c(0, 2, -1)))

trueDTRmeans
```

### Marginal Model, adjust for baseline covariates

```{r}
fitmm.adj <- geepack::geeglm(Y2 ~ A1 + A1:A2, data = autism.mm, id = ID, weights = w)
summary(fitmm.adj)

# Careful we specify contrasts correctly
estimate(fitmm.adj,
         rbind(
           # These statements get the mean under each AI
           # Remember the coefficients are positional: the first is the intercept, the second is the variable in the blank, the third is a1a2.
           "Mean Y: AI#1 (JASP+EMT,INTENSFY)" = c(1, 1, 1),
           "Mean Y: AI#2 (JASP+EMT, Add SGD)" = c(1, 1, -1),
           "Mean Y: AI#3 (JASP+EMT+SGD, ...)" = c(1, -1, 0),
           "Diff: (JASP+EMT,Add SGD) - (JASP+EMT+SGD,...)" = c(0, 2, -1)))

trueDTRmeans
```

# Module 5 / DEMO 3

### Moderators of Second-stage TX effects

First-stage effects will be biased Why do we subset to only non-responders? We loose power adjusting for baseline and first-stage?

```{r}
model2s.c <- autism.nr.c %>% geeglm(Y2 ~ O11.c + O12.c + A2*(O21.c + O22.c),
                                id = ID, data = .)
summary(model2s.c) #gives robust SE
```

### Interaction Plot 2nd Stage for A1==1

```{r}
library(emmeans)

model2s <- autism.nr.c %>% geeglm(Y2 ~ O11.c + O12.c + A2*(O21 + O22),
                                id = ID, data = .)

grid <- emmeans::ref_grid(model2s, cov.keep = c("A2", "O21"))

em2 <- emmeans::emmeans(grid, ~ A2) # This does not give a weighted grid!!!! use ref_grid or weights = "proportional"
summary(em2)
plot(em2)


emmeans::emmip(grid, A2 ~ O21) # use grid object 


```

### Q-learning

control for optimal second-stage tactic

```{r}
# Using full dataset, centered first, then filtered to Rerand
autism.c.nr <- autism.c %>% filter(Rerand == 1)

model2s <-  autism.c.nr %>%
  geeglm(Y2 ~ O11.c + O12.c + A2*(O21 + O22), 
         id = ID, data = .)

autism.c.nr.alt <- autism.c.nr %>% mutate(A2 = -1 * A2)
autism.c.nr.opt <- autism.c.nr

autism.c.nr.alt$Y2 <- predict(model2s, newdata = autism.c.nr.alt) # predict response on counterfactual a2

for (i in 1:nrow(autism.c.nr.alt)) {
  if (autism.c.nr.alt$Y2[i] > autism.c.nr.opt$Y2[i]) {
    autism.c.nr.opt[i, ] <- autism.c.nr.alt[i, ] # assign optimal A2
  }
}

# Combine Optimal non-responders rest
autism.c.opt <- autism.c %>% filter(Rerand == 0) %>%
  bind_rows(autism.c.nr.opt)

# Fit first-stage model controlling for second-stage
model1s <- autism.c.opt %>% geeglm(Y2 ~ O11 + O12 + A1*(O11 + O12), id = ID, data = .)
summary(model1s)


```

### Interaction plot 1st Stage; controlling for optimal second stage

```{r}
grid <- emmeans::ref_grid(model1s, cov.keep = c("A1","O11"))

em1 <- emmeans::emmeans(model1s, ~ A1, weights = "proportional")
summary(em1)
plot(em1)

emmeans::emmip(grid, A1 ~ O11) 
```

TO DO: calculate mean of more deeply tailored AI

### qlaci

Need to implement qlaci package. Are we using?? I think there are some errors in the Q-learning demo...

### Adjust for Time-varying Confounders (conditional model)

Two-stage regression algorithm. SE errors need to be adjusted

```{r}

# Response model
fitR <- autism.c %>% glm(R ~ O11.c + O12.c + A1,
                       family = binomial, data = .)
summary(fitR)

# O21 model
fitO21 <- autism.c %>% lm(O21 ~ O11.c + O12.c + A1 + R, data = .)
summary(fitO21)

# O22 model
fitO22 <- autism.c %>% lm(O22 ~ O11.c + O12.c + A1 + R + O21, data = .)
summary(fitO22)

# Add residuals to df
autism.c$R.resid <- residuals(fitR, "response")
autism.c$O21.resid <- residuals(fitO21, "response")
autism.c$O22.resid <- residuals(fitO22, "response")

# Fit main model
fitCM <- autism.c %>% geeglm(Y2 ~ A1*R.resid + Rerand:A2, data = ., id = ID)

summary(fitCM)


```

### Moderator Analysis

```{r}

# Fit main model
fitCM.mod <- autism.c %>% geeglm(Y2 ~ A1*(R.resid + O21.resid + O22.resid) + Rerand:(A2/(O21.resid + O22.resid)), data = ., id = ID)

summary(fitCM.mod)

```
